### 2018-03-16 paper/paper.tex
start paper template

\documentclass[letterpaper]{article}

\input{paper.sty}
% Load nips_2017 last to override any changes I didn't mean to make.
\usepackage[final]{nips_2017}

\title{Generating haikus with an LSTM network}

\author{
    Austin Gill\\
    Department of Mathematics and Computer Science\\
    South Dakota School of Mines and Technology\\
    Rapid City SD, 57701\\
    \texttt{\href{mailto:austin.gill@mines.sdsmt.edu}{austin.gill@mines.sdsmt.edu}} \\
}

\begin{document}
\maketitle

\begin{abstract}
    Here's an abstract.
\end{abstract}

\section{Introduction}
    Here's the introduction.

\section{Stuff}
    Here's some stuff you didn't know!

\section{Conclusion}
    Conclude something\todo{Something useful!}

\end{document}

### 2018-03-18 data/scrape_haikus.py
Add tweet generation script
* It's still training as I write this message. It's generating stuff
like this:

> this will be our last chance there will never be another opportunity
> march that is the best the secret a state of the dems the best that the
> many that the world that the world that is the best that the republican
> party is a great people who wants to see the new season of the debate
> than the state of the dems the best people that the world that is the
> best the world that we have to see the most incredible that the world
> that we have to be the republican party with the debat

@@ -1,4 +1,9 @@
 #!/usr/bin/env python3
+
+"""
+    Downloads haikus from different collections on the internet.
+"""
+
 import string
 from requests_html import HTMLSession
 

### 2018-03-18 experiments/dataset.py

@@ -1,0 +1,61 @@
+import json
+import re
+import string
+from zipfile import ZipFile
+
+
+def remove_urls(vTEXT):
+    """
+        Remove URLS from a given string.
+    """
+    vTEXT = re.sub(r'(https|http)?:\/\/(\w|\.|\/|\?|\=|\&|\%)*\b', '', vTEXT, flags=re.MULTILINE)
+    return vTEXT
+
+
+# Redefine what 'alphabet' means so we don't filter out spaces and hashtags.
+ALPHABET = frozenset(string.ascii_lowercase + ' ' + '#')
+
+
+def preprocess(text, use_ascii=True):
+    """
+        Preprocess text. Converts to lowercase and filters non-alphabetic characters.
+        Defaults to defining alphabetic characters as ascii-alphabetic.
+
+        Examples:
+        >>> text = 'ABC.,#'
+        >>> ''.join(preprocess(text))
+        'abc'
+        >>> text = 'ÈÆÖÉEAEOE,.%'
+        >>> ''.join(preprocess(text, use_ascii=False))
+        'èæöéeaeoe'
+    """
+    if use_ascii:
+        return filter(ALPHABET.__contains__, text.lower())
+    return filter(str.isalpha, text.lower())
+
+
+def load_dataset(files):
+    """
+        Parses the tweets out of the zipped dataset. Returns a generator of preprocessed tweets.
+    """
+    dataset = []
+
+    for archive in files:
+        with ZipFile(archive) as z:
+            # ZipFile's are collections of zipped files.
+            for name in z.namelist():
+                with z.open(name) as f:
+                    print(f'reading {name}')
+                    dataset += json.load(f)
+
+    # Filter out the retweets
+    dataset = filter(lambda x: x['is_retweet'] is False, dataset)
+    # Get only the text
+    dataset = (t['text']for t in dataset)
+    # Remove URLs from the tweets.
+    # TODO: Consider removing username mentions?
+    dataset = (remove_urls(t) for t in dataset)
+    # Preprocess each tweet, filtering out nonascii alphabetic
+    dataset = (''.join(preprocess(t)) for t in dataset)
+
+    return dataset

### 2018-03-18 experiments/tweets.py

@@ -1,9 +1,11 @@
 #!/usr/bin/env python3
 
+import glob
 import random
 import sys
 
 import numpy as np
+from dataset import load_dataset
 
 # Hack to keep keras from allocating the whole damn gpu.
 import tensorflow as tf
@@ -19,10 +21,6 @@ from keras.callbacks import LambdaCallback
 from keras.layers.core import Dense, Dropout
 from keras.layers.recurrent import LSTM
 from keras.models import Sequential
-from keras.preprocessing.sequence import pad_sequences
-
-
-
 
 '''
     Example script to generate haiku Text.
@@ -32,26 +30,27 @@ from keras.preprocessing.sequence import pad_sequences
     has at least ~100k characters. ~1M is better.
 '''
 
-path = "../data/haikus.txt"
-text = open(path).read().lower()
-print('corpus length:', len(text))
+text = load_dataset(glob.glob('../data/trump_tweet_data_archive/condensed_*.json.zip'))
+# turn the list of tweets into one big corpus
+text = ' '.join(text)
+print(f'corpus length: {len(text)}')
 
 chars = sorted(list(set(text)))
-print('total chars:', len(chars))
+print(f'total chars: {len(chars)}')
 char_indices = dict((c, i) for i, c in enumerate(chars))
 indices_char = dict((i, c) for i, c in enumerate(chars))
 
 # cut the text in semi-redundant sequences of maxlen characters
-maxlen = 45
+maxlen = 120
 step = 3
 sentences = []
 next_chars = []
 for i in range(0, len(text) - maxlen, step):
     sentences.append(text[i: i + maxlen])
     next_chars.append(text[i + maxlen])
-print('nb sequences:', len(sentences))
+print(f'sequences: {len(sentences)}')
 
-print('Vectorization...')
+print('Vectorizing input...')
 X = np.zeros((len(sentences), maxlen, len(chars)), dtype=np.bool)
 y = np.zeros((len(sentences), len(chars)), dtype=np.bool)
 for i, sentence in enumerate(sentences):
@@ -60,9 +59,11 @@ for i, sentence in enumerate(sentences):
     y[i, char_indices[next_chars[i]]] = 1
 
 print(f'X shape: {X.shape}, Y shape: {y.shape}')
+print(f'X size: {sys.getsizeof(X) * 0.000001 :.3f} MB')
+print(f'y size: {sys.getsizeof(y) * 0.000001 :.3f} MB')
 
 # build the model: single LSTM
-print('Build model...')
+print('Building model...')
[…]

### 2018-03-18 experiments/dataset.py
Add verbosity to load_dataset()

@@ -34,7 +34,7 @@ def preprocess(text, use_ascii=True):
     return filter(str.isalpha, text.lower())
 
 
-def load_dataset(files):
+def load_dataset(files, verbose=True):
     """
         Parses the tweets out of the zipped dataset. Returns a generator of preprocessed tweets.
     """
@@ -45,7 +45,8 @@ def load_dataset(files):
             # ZipFile's are collections of zipped files.
             for name in z.namelist():
                 with z.open(name) as f:
-                    print(f'reading {name}')
+                    if verbose:
+                        print(f'reading {name}')
                     dataset += json.load(f)
 
     # Filter out the retweets

### 2018-03-18 experiments/tweets.py

@@ -5,22 +5,22 @@ import random
 import sys
 
 import numpy as np
+import tensorflow as tf
+from keras.backend.tensorflow_backend import set_session
+from keras.callbacks import LambdaCallback
+from keras.layers.core import Dense, Dropout
+from keras.layers.recurrent import LSTM
+from keras.models import Sequential
+
 from dataset import load_dataset
 
 # Hack to keep keras from allocating the whole damn gpu.
-import tensorflow as tf
-from keras.backend.tensorflow_backend import set_session
-
 config = tf.ConfigProto()
 config.gpu_options.allow_growth = True
 config.gpu_options.visible_device_list = "0"
 #session = tf.Session(config=config)
 set_session(tf.Session(config=config))
 
-from keras.callbacks import LambdaCallback
-from keras.layers.core import Dense, Dropout
-from keras.layers.recurrent import LSTM
-from keras.models import Sequential
 
 '''
     Example script to generate haiku Text.

### 2018-03-18 experiments/model.py
Attempt to make code better. Still some things to figure out

@@ -1,0 +1,233 @@
+#!/usr/bin/env python3
+import random
+import sys
+from datetime import datetime
+from glob import glob
+
+import matplotlib.pyplot as plt
+import numpy as np
+import tensorflow as tf
+from keras.backend.tensorflow_backend import set_session
+from keras.callbacks import LambdaCallback
+from keras.layers.core import Dense, Dropout
+from keras.layers.recurrent import LSTM
+from keras.models import Sequential, model_from_json
+
+from dataset import load_dataset
+
+# Hack to keep keras from allocating the whole damn gpu.
+config = tf.ConfigProto()
+config.gpu_options.allow_growth = True
+config.gpu_options.visible_device_list = "0"
+#session = tf.Session(config=config)
+set_session(tf.Session(config=config))
+
+
+def chunk_text(corpus, length, step):
+    """
+        Cut the given corpus into a series of offset sequences.
+    """
+
+    # A list of sequences
+    sequences = []
+    # For each sequence, record the next character in the corpus
+    next_chars = []
+    for i in range(0, len(corpus) - length, step):
+        sequences.append(corpus[i:i + length])
+        next_chars.append(corpus[i + length])
+    return sequences, next_chars
+
+
+def vectorize_data(corpus, sequences, next_chars, seq_length, num_chars, char_to_indices, verbose):
+    """
+        Given the textual dataset, vectorize it.
+    """
+    if verbose:
+        print('Vectorizing sequences...')
+    X = np.zeros((len(sequences), seq_length, num_chars), dtype=np.bool)
+    y = np.zeros((len(sequences), num_chars), dtype=np.bool)
+    for i, seq in enumerate(sequences):
+        for t, char in enumerate(seq):
+            X[i, t, char_to_indices[char]] = 1
+        y[i, char_to_indices[next_chars[i]]] = 1
+
+    return X, y
+
+
+def build_model(seq_length, num_chars, verbose):
+    """
+        Builds an LSTM model for generating text. One layer should be enough.
+    """
+    if verbose:
+        print('Building model...')
+    model = Sequential()
+    # TODO: Model size?
+    model.add(LSTM(256, input_shape=(seq_length, num_chars)))
+    # TODO: Regularization?
+    # model.add(Dropout(0.2))
+    model.add(Dense(num_chars, activation='softmax'))
+
+    # TODO: Try different loss functions, optimizers, and metrics
[…]

### 2018-03-19 experiments/model.py
Fix syntax errors.

@@ -132,7 +132,7 @@ def save_model(model, verbose, filename=None):
         print(f'saving model weights to {filename}.h5')
 
     with open(f'{filename}.json', 'w') as json:
-        json.write(model.to_json)
+        json.write(model.to_json())
     model.save_weights(f'{filename}.h5')
 
 
@@ -146,7 +146,7 @@ def load_model(base_filename, verbose):
         print(f'Loading model weights from {base_filename}.h5')
 
     with open(f'{base_filename}.json') as json:
-        model = model_from_json(json)
+        model = model_from_json(json.read())
     with open(f'{base_filename}.h5') as hdf5:
         model.load_weights(f'{base_filename}.h5')
     return model
@@ -170,7 +170,7 @@ def generate_sequence(model, corpus, seed, length, diversities,
             print(f'Generated:')
 
         for _ in range(length):
-            X_pred = np.zeroes((1, seq_length, num_chars))
+            X_pred = np.zeros((1, seq_length, num_chars))
             for t, c in enumerate(sequence):
                 X_pred[0, t, char_to_indices[c]] = 1
 
@@ -216,10 +216,12 @@ def main(verbose):
 
     model = build_model(LEN, num_chars, verbose)
     h = train_model(model, X, y, verbose=verbose)
-    save_model(model, verbose)
+    save_model(model, verbose, filename='latest')
 
-    # TODO: These will probably have to be LEN chars long
-    seeds = ['republicans', 'hate', 'fake news', 'golf']
+    # model = load_model('latest', verbose)
+
+    # TODO: These will probably have to be LEN chars long?
+    seeds = ['republicans', 'democrats', 'hate', 'fake news', 'golf']
     generated = []
     for seed in seeds:
         generate_sequence(model, corpus, seed, 200, [0.05, 0.1, 0.2, 0.5], LEN, num_chars,

### 2018-03-19 experiments/tweets.py

@@ -23,7 +23,7 @@ set_session(tf.Session(config=config))
 
 
 '''
-    Example script to generate haiku Text.
+    Example script to generate tweets.
     It is recommended to run this script on GPU, as recurrent
     networks are quite computationally intensive.
     If you try this script on new data, make sure your corpus

### 2018-03-19 experiments/dataset.py
Remove @mentions from dataset

@@ -11,8 +11,16 @@ def remove_urls(vTEXT):
     vTEXT = re.sub(r'(https|http)?:\/\/(\w|\.|\/|\?|\=|\&|\%)*\b', '', vTEXT, flags=re.MULTILINE)
     return vTEXT
 
+def remove_mentions(vTEXT):
+    """
+        Remove @mentions from a given string.
+    """
+    vTEXT = re.sub(r'@\w+', '', vTEXT, flags=re.MULTILINE)
+    return vTEXT
+
 
 # Redefine what 'alphabet' means so we don't filter out spaces and hashtags.
+# TODO: Consider leaving '@'s in
 ALPHABET = frozenset(string.ascii_lowercase + ' ' + '#')
 
 
@@ -53,8 +61,9 @@ def load_dataset(files, verbose=True):
     dataset = filter(lambda x: x['is_retweet'] is False, dataset)
     # Get only the text
     dataset = (t['text']for t in dataset)
+    # Remove @mentions from the tweets.
+    dataset = (remove_mentions(t) for t in dataset)
     # Remove URLs from the tweets.
-    # TODO: Consider removing username mentions?
     dataset = (remove_urls(t) for t in dataset)
     # Preprocess each tweet, filtering out nonascii alphabetic
     dataset = (''.join(preprocess(t)) for t in dataset)

### 2018-03-19 experiments/model.py
Add validation data to training step

@@ -8,7 +8,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import tensorflow as tf
 from keras.backend.tensorflow_backend import set_session
-from keras.callbacks import LambdaCallback
 from keras.layers.core import Dense, Dropout
 from keras.layers.recurrent import LSTM
 from keras.models import Sequential, model_from_json
@@ -61,7 +60,7 @@ def build_model(seq_length, num_chars, verbose):
     if verbose:
         print('Building model...')
     model = Sequential()
-    # TODO: Model size?
+    # TODO: Model size/depth?
     model.add(LSTM(256, input_shape=(seq_length, num_chars)))
     # TODO: Regularization?
     # model.add(Dropout(0.2))
@@ -84,7 +83,7 @@ def sample_predictions(preds, temperature=1.0):
     return np.argmax(probas)
 
 
-def train_model(model, X, y, val_X=None, val_y=None, verbose=True):
+def train_model(model, X, y, X_val=None, y_val=None, verbose=True):
     """
         Trains the given model. Will use validation data if given.
 
@@ -93,11 +92,11 @@ def train_model(model, X, y, val_X=None, val_y=None, verbose=True):
     BATCH_SIZE = 128
     EPOCHS = 50
 
-    if val_X and val_y:
+    if X_val and y_val:
         h = model.fit(X, y,
                       batch_size=BATCH_SIZE,
                       epochs=EPOCHS,
-                      validation_data=(val_X, val_y))
+                      validation_data=(X_val, y_val))
     else:
         h = model.fit(X, y,
                       batch_size=BATCH_SIZE,
@@ -108,7 +107,7 @@ def train_model(model, X, y, val_X=None, val_y=None, verbose=True):
 
     plt.plot(h.history['loss'], label='training loss')
 
-    if val_X and val_y:
+    if X_val and y_val:
         plt.plot(h.history['val_loss'], label='validation loss')
     plt.ylabel('loss')
     plt.xlabel('epochs')
@@ -147,8 +146,7 @@ def load_model(base_filename, verbose):
 
     with open(f'{base_filename}.json') as json:
         model = model_from_json(json.read())
-    with open(f'{base_filename}.h5') as hdf5:
-        model.load_weights(f'{base_filename}.h5')
+    model.load_weights(f'{base_filename}.h5')
[…]

### 2018-03-19 experiments/model.py
Use longer seeds taken from the corpus

@@ -230,14 +230,16 @@ def main(verbose):
         print(f'X_val size: {sys.getsizeof(X_val) * 0.000001 :.3f} MB')
         print(f'y_val size: {sys.getsizeof(y_val) * 0.000001 :.3f} MB')
 
-    model = build_model(LEN, num_chars, verbose)
-    h = train_model(model, X_train, y_train, X_val=X_val, y_val=y_val, verbose=verbose)
-    save_model(model, verbose, filename='latest')
+    # model = build_model(LEN, num_chars, verbose)
+    # h = train_model(model, X_train, y_train, X_val=X_val, y_val=y_val, verbose=verbose)
+    # save_model(model, verbose, filename='latest')
 
-    # model = load_model('latest', verbose)
+    # TODO: train/load as argument
+    model = load_model('latest', verbose)
 
-    # TODO: These will probably have to be LEN chars long?
-    seeds = ['republicans', 'democrats', 'hate', 'fake news', 'golf']
+    # Generate text from seeds randomly taken from the corpus
+    indices = [random.randint(0, len(corpus) - LEN - 1) for _ in range(10)]
+    seeds = [corpus[i: i + LEN] for i in indices]
     generated = []
     for seed in seeds:
         generate_sequence(model, corpus, seed, 200, [0.05, 0.1, 0.2, 0.5], LEN, num_chars,

### 2018-03-19 experiments/model.py
Implement training as an argument

@@ -61,6 +61,7 @@ def build_model(seq_length, num_chars, verbose):
         print('Building model...')
     model = Sequential()
     # TODO: Model size/depth?
+    # TODO: LSTM options?
     model.add(LSTM(256, input_shape=(seq_length, num_chars)))
     # TODO: Regularization?
     # model.add(Dropout(0.2))
@@ -183,7 +184,7 @@ def generate_sequence(model, corpus, seed, length, diversities,
             print(generated)
 
 
-def main(verbose):
+def main(train, verbose):
     corpus = load_dataset(glob('../data/trump_tweet_data_archive/condensed_*.json.zip'), verbose)
     corpus = ' '.join(corpus)
     characters = sorted(list(set(corpus)))
@@ -204,7 +205,6 @@ def main(verbose):
         print(f'num characters: {num_chars}')
         print(f'number of sequences: {len(sequences)}')
 
-    # TODO: Split into training and validation data
     X, y = vectorize_data(corpus, sequences, next_chars, LEN, num_chars, char_to_indices, verbose)
 
     if verbose:
@@ -230,12 +230,12 @@ def main(verbose):
         print(f'X_val size: {sys.getsizeof(X_val) * 0.000001 :.3f} MB')
         print(f'y_val size: {sys.getsizeof(y_val) * 0.000001 :.3f} MB')
 
-    # model = build_model(LEN, num_chars, verbose)
-    # h = train_model(model, X_train, y_train, X_val=X_val, y_val=y_val, verbose=verbose)
-    # save_model(model, verbose, filename='latest')
-
-    # TODO: train/load as argument
-    model = load_model('latest', verbose)
+    if train:
+        model = build_model(LEN, num_chars, verbose)
+        h = train_model(model, X_train, y_train, X_val=X_val, y_val=y_val, verbose=verbose)
+        save_model(model, verbose, filename='latest')
+    else:
+        model = load_model('latest', verbose)
 
     # Generate text from seeds randomly taken from the corpus
     indices = [random.randint(0, len(corpus) - LEN - 1) for _ in range(10)]
@@ -250,4 +250,5 @@ def main(verbose):
 
 
 if __name__ == '__main__':
-    main(verbose=True)
+    train = '--train' in sys.argv
+    main(verbose=True, train=train)

### 2018-03-19 experiments/model.py
Eliminate some cruft

@@ -206,12 +206,6 @@ def main(train, verbose):
         print(f'number of sequences: {len(sequences)}')
 
     X, y = vectorize_data(corpus, sequences, next_chars, LEN, num_chars, char_to_indices, verbose)
-
-    if verbose:
-        print(f'X shape: {X.shape}, Y shape: {y.shape}')
-        print(f'X size: {sys.getsizeof(X) * 0.000001 :.3f} MB')
-        print(f'y size: {sys.getsizeof(y) * 0.000001 :.3f} MB')
-
     n = len(X)
     # Use 20% validation data
     num_val = int(0.2 * n)
@@ -223,12 +217,9 @@ def main(train, verbose):
 
     if verbose:
         print(f'Number validation samples: {num_val}')
-        print(f'X_train shape: {X_train.shape}, Y_train shape: {y_train.shape}')
-        print(f'X_train size: {sys.getsizeof(X_train) * 0.000001 :.3f} MB')
-        print(f'y_train size: {sys.getsizeof(y_train) * 0.000001 :.3f} MB')
-        print(f'X_val shape: {X_val.shape}, Y_val shape: {y_val.shape}')
-        print(f'X_val size: {sys.getsizeof(X_val) * 0.000001 :.3f} MB')
-        print(f'y_val size: {sys.getsizeof(y_val) * 0.000001 :.3f} MB')
+        print(f'X shape: {X.shape}, Y shape: {y.shape}')
+        print(f'X size: {sys.getsizeof(X) * 0.000001 :.3f} MB')
+        print(f'y size: {sys.getsizeof(y) * 0.000001 :.3f} MB')
 
     if train:
         model = build_model(LEN, num_chars, verbose)

### 2018-03-20 experiments/dataset.py
Add username @mentions back and fix validation

@@ -19,9 +19,8 @@ def remove_mentions(vTEXT):
     return vTEXT
 
 
-# Redefine what 'alphabet' means so we don't filter out spaces and hashtags.
-# TODO: Consider leaving '@'s in
-ALPHABET = frozenset(string.ascii_lowercase + ' ' + '#')
+# Redefine what 'alphabet' means so we don't filter out spaces, @mentions, and hashtags.
+ALPHABET = frozenset(string.ascii_lowercase + ' ' + '#' + '@')
 
 
 def preprocess(text, use_ascii=True):
@@ -61,8 +60,8 @@ def load_dataset(files, verbose=True):
     dataset = filter(lambda x: x['is_retweet'] is False, dataset)
     # Get only the text
     dataset = (t['text']for t in dataset)
-    # Remove @mentions from the tweets.
-    dataset = (remove_mentions(t) for t in dataset)
+    # # Remove @mentions from the tweets.
+    # dataset = (remove_mentions(t) for t in dataset)
     # Remove URLs from the tweets.
     dataset = (remove_urls(t) for t in dataset)
     # Preprocess each tweet, filtering out nonascii alphabetic

### 2018-03-20 experiments/model.py

@@ -8,8 +8,9 @@ import matplotlib.pyplot as plt
 import numpy as np
 import tensorflow as tf
 from keras.backend.tensorflow_backend import set_session
-from keras.layers.core import Dense, Dropout
-from keras.layers.recurrent import LSTM
+from keras.layers import Dense, Dropout
+# CuDNNLSTM is a CUDA accelerated LSTM layer
+from keras.layers import CuDNNLSTM
 from keras.models import Sequential, model_from_json
 
 from dataset import load_dataset
@@ -62,9 +63,9 @@ def build_model(seq_length, num_chars, verbose):
     model = Sequential()
     # TODO: Model size/depth?
     # TODO: LSTM options?
-    model.add(LSTM(256, input_shape=(seq_length, num_chars)))
+    model.add(CuDNNLSTM(units=512, input_shape=(seq_length, num_chars), unit_forget_bias=True))
+    model.add(Dropout(0.2))
     # TODO: Regularization?
-    # model.add(Dropout(0.2))
     model.add(Dense(num_chars, activation='softmax'))
 
     # TODO: Try different loss functions, optimizers, and metrics
@@ -93,7 +94,9 @@ def train_model(model, X, y, X_val=None, y_val=None, verbose=True):
     BATCH_SIZE = 128
     EPOCHS = 50
 
-    if X_val and y_val:
+    if X_val is not None and y_val is not None:
+        if verbose:
+            print(f'Training with {len(X_val)} validation samples')
         h = model.fit(X, y,
                       batch_size=BATCH_SIZE,
                       epochs=EPOCHS,
@@ -108,7 +111,7 @@ def train_model(model, X, y, X_val=None, y_val=None, verbose=True):
 
     plt.plot(h.history['loss'], label='training loss')
 
-    if X_val and y_val:
+    if X_val is not None and y_val is not None:
         plt.plot(h.history['val_loss'], label='validation loss')
     plt.ylabel('loss')
     plt.xlabel('epochs')
@@ -156,8 +159,6 @@ def generate_sequence(model, corpus, seed, length, diversities,
     """
         Given a model, generate a text given some seed.
     """
-    # TODO: Does len(seed) == LEN?
-
     for diversity in diversities:
         generated = ''
         sequence = seed
@@ -209,11 +210,11 @@ def main(train, verbose):
     n = len(X)
     # Use 20% validation data
     num_val = int(0.2 * n)
-    X_val = X[num_val:]
[…]

### 2018-03-20 experiments/model.py
Swap out CUDA LSTM layers with CPU LSTM layers

@@ -10,7 +10,9 @@ import tensorflow as tf
 from keras.backend.tensorflow_backend import set_session
 from keras.layers import Dense, Dropout
 # CuDNNLSTM is a CUDA accelerated LSTM layer
-from keras.layers import CuDNNLSTM
+# (Not so good for running (even trained) models on a laptop)
+# from keras.layers import CuDNNLSTM
+from keras.layers import LSTM
 from keras.models import Sequential, model_from_json
 
 from dataset import load_dataset
@@ -63,7 +65,7 @@ def build_model(seq_length, num_chars, verbose):
     model = Sequential()
     # TODO: Model size/depth?
     # TODO: LSTM options?
-    model.add(CuDNNLSTM(units=512, input_shape=(seq_length, num_chars), unit_forget_bias=True))
+    model.add(LSTM(units=512, input_shape=(seq_length, num_chars), unit_forget_bias=True))
     model.add(Dropout(0.2))
     # TODO: Regularization?
     model.add(Dense(num_chars, activation='softmax'))
@@ -225,9 +227,9 @@ def main(train, verbose):
     if train:
         model = build_model(LEN, num_chars, verbose)
         h = train_model(model, X_train, y_train, X_val=X_val, y_val=y_val, verbose=verbose)
-        save_model(model, verbose, filename='512dropout')
+        save_model(model, verbose, filename='models/512dropout')
     else:
-        model = load_model('latest', verbose)
+        model = load_model('models/512dropout', verbose)
 
     # Generate text from seeds randomly taken from the corpus
     indices = [random.randint(0, len(corpus) - LEN - 1) for _ in range(10)]

### 2018-03-21 experiments/tweets.py
Remove outdated script

@@ -1,125 +1,0 @@
-#!/usr/bin/env python3
-
-import glob
-import random
-import sys
-
-import numpy as np
-import tensorflow as tf
-from keras.backend.tensorflow_backend import set_session
-from keras.callbacks import LambdaCallback
-from keras.layers.core import Dense, Dropout
-from keras.layers.recurrent import LSTM
-from keras.models import Sequential
-
-from dataset import load_dataset
-
-# Hack to keep keras from allocating the whole damn gpu.
-config = tf.ConfigProto()
-config.gpu_options.allow_growth = True
-config.gpu_options.visible_device_list = "0"
-#session = tf.Session(config=config)
-set_session(tf.Session(config=config))
-
-
-'''
-    Example script to generate tweets.
-    It is recommended to run this script on GPU, as recurrent
-    networks are quite computationally intensive.
-    If you try this script on new data, make sure your corpus
-    has at least ~100k characters. ~1M is better.
-'''
-
-text = load_dataset(glob.glob('../data/trump_tweet_data_archive/condensed_*.json.zip'))
-# turn the list of tweets into one big corpus
-text = ' '.join(text)
-print(f'corpus length: {len(text)}')
-
-chars = sorted(list(set(text)))
-print(f'total chars: {len(chars)}')
-char_indices = dict((c, i) for i, c in enumerate(chars))
-indices_char = dict((i, c) for i, c in enumerate(chars))
-
-# cut the text in semi-redundant sequences of maxlen characters
-maxlen = 120
-step = 3
-sentences = []
-next_chars = []
-for i in range(0, len(text) - maxlen, step):
-    sentences.append(text[i: i + maxlen])
-    next_chars.append(text[i + maxlen])
-print(f'sequences: {len(sentences)}')
-
-print('Vectorizing input...')
-X = np.zeros((len(sentences), maxlen, len(chars)), dtype=np.bool)
-y = np.zeros((len(sentences), len(chars)), dtype=np.bool)
-for i, sentence in enumerate(sentences):
-    for t, char in enumerate(sentence):
-        X[i, t, char_indices[char]] = 1
-    y[i, char_indices[next_chars[i]]] = 1
-
-print(f'X shape: {X.shape}, Y shape: {y.shape}')
-print(f'X size: {sys.getsizeof(X) * 0.000001 :.3f} MB')
-print(f'y size: {sys.getsizeof(y) * 0.000001 :.3f} MB')
-
-# build the model: single LSTM
-print('Building model...')
-model = Sequential()
[…]

### 2018-03-22 experiments/model.py
Add TODO items, shuffle training data

@@ -52,8 +52,10 @@ def vectorize_data(corpus, sequences, next_chars, seq_length, num_chars, char_to
         for t, char in enumerate(seq):
             X[i, t, char_to_indices[char]] = 1
         y[i, char_to_indices[next_chars[i]]] = 1
-
-    return X, y
+    # An array of indices
+    I = np.arange(X.shape[0])
+    np.random.shuffle(I)
+    return X[I], y[I]
 
 
 def build_model(seq_length, num_chars, verbose):
@@ -94,7 +96,7 @@ def train_model(model, X, y, X_val=None, y_val=None, verbose=True):
         Will also save loss plots.
     """
     BATCH_SIZE = 128
-    EPOCHS = 50
+    EPOCHS = 30
 
     if X_val is not None and y_val is not None:
         if verbose:
@@ -197,7 +199,7 @@ def main(train, verbose):
     indices_to_char = dict((i, c) for i, c in enumerate(characters))
 
     # The length of the sequences
-    LEN = 60
+    LEN = 25
     # How the far apart the sequences are spaced
     STEP = 3
 
@@ -227,9 +229,9 @@ def main(train, verbose):
     if train:
         model = build_model(LEN, num_chars, verbose)
         h = train_model(model, X_train, y_train, X_val=X_val, y_val=y_val, verbose=verbose)
-        save_model(model, verbose, filename='models/512dropout')
+        save_model(model, verbose, filename='512dropout')
     else:
-        model = load_model('models/512dropout', verbose)
+        model = load_model('512dropout', verbose)
 
     # Generate text from seeds randomly taken from the corpus
     indices = [random.randint(0, len(corpus) - LEN - 1) for _ in range(10)]

### 2018-04-16 experiments/generate.py
Do not output the seed along with the generated text

@@ -34,7 +34,7 @@ def generate_sequence(model, seed, length, diversities,
     for diversity in diversities:
         generated = ''
         sequence = seed
-        generated += seed
+        # generated += seed
 
         if verbose:
             print('='*10)

### 2018-04-16 experiments/model.py
Try a deeper layer. Results seem pretty good.

@@ -20,9 +20,9 @@ GEN_LEN = 280
 # Default is 32.
 BATCH_SIZE = 128
 # Number of epochs to train for. RMSprop is *much* faster than SGD and only needs 5-10 epochs
-EPOCHS = 20
+EPOCHS = 50
 # Extensionless filename to save model configuration, weights, plots, and generated text to
-BASENAME = 'models/128-128-rms-batchnorm-20'
+BASENAME = 'models/64x6-rms-batchnorm-50'
 
 
 def build_model(seq_length, num_chars, verbose):
@@ -32,9 +32,17 @@ def build_model(seq_length, num_chars, verbose):
     if verbose:
         print('Building model...')
     model = Sequential()
-    model.add(LSTM(units=128, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
+    model.add(LSTM(units=64, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
     model.add(BatchNormalization())
-    model.add(LSTM(units=128, unit_forget_bias=True))
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, unit_forget_bias=True))
     model.add(BatchNormalization())
     model.add(Dense(num_chars, activation='softmax'))
 

### 2018-04-17 experiments/textgenrnn/textgenrnn/model.py
move textgenrnn

@@ -1,0 +1,87 @@
+from keras.optimizers import RMSprop
+from keras.layers import Input, Embedding, Dense, LSTM, Bidirectional
+from keras.layers import CuDNNLSTM, concatenate, Reshape
+from keras.models import Model
+from keras import backend as K
+from .AttentionWeightedAverage import AttentionWeightedAverage
+
+
+def textgenrnn_model(num_classes, cfg, context_size=None,
+                     weights_path=None,
+                     optimizer=RMSprop(lr=4e-3, rho=0.99)):
+    '''
+    Builds the model architecture for textgenrnn and
+    loads the specified weights for the model.
+    '''
+
+    input = Input(shape=(cfg['max_length'],), name='input')
+    embedded = Embedding(num_classes, cfg['dim_embeddings'],
+                         input_length=cfg['max_length'],
+                         name='embedding')(input)
+
+    rnn_layer_list = []
+    for i in range(cfg['rnn_layers']):
+        prev_layer = embedded if i is 0 else rnn_layer_list[-1]
+        rnn_layer_list.append(new_rnn(cfg, i+1)(prev_layer))
+
+    seq_concat = concatenate([embedded] + rnn_layer_list, name='rnn_concat')
+    attention = AttentionWeightedAverage(name='attention')(seq_concat)
+    output = Dense(num_classes, name='output', activation='softmax')(attention)
+
+    if context_size is None:
+        model = Model(inputs=[input], outputs=[output])
+        if weights_path is not None:
+            model.load_weights(weights_path, by_name=True)
+        model.compile(loss='categorical_crossentropy', optimizer=optimizer)
+
+    else:
+        context_input = Input(
+            shape=(context_size,), name='context_input')
+        context_reshape = Reshape((context_size,),
+                                  name='context_reshape')(context_input)
+        merged = concatenate([attention, context_reshape], name='concat')
+        main_output = Dense(num_classes, name='context_output',
+                            activation='softmax')(merged)
+
+        model = Model(inputs=[input, context_input],
+                      outputs=[main_output, output])
+        if weights_path is not None:
+            model.load_weights(weights_path, by_name=True)
+        model.compile(loss='categorical_crossentropy', optimizer=optimizer,
+                      loss_weights=[0.8, 0.2])
+
+    return model
+
+
+'''
+Create a new LSTM layer per parameters. Unfortunately,
+each combination of parameters must be hardcoded.
+
[…]

### 2018-04-17 experiments/textgenrnn/textgenrnn/textgenrnn.py

@@ -1,0 +1,420 @@
+from keras.callbacks import LearningRateScheduler, Callback
+from keras.models import Model, load_model
+from keras.preprocessing import sequence
+from keras.preprocessing.text import Tokenizer, text_to_word_sequence
+from keras import backend as K
+from sklearn.preprocessing import LabelBinarizer
+import numpy as np
+import json
+import h5py
+from pkg_resources import resource_filename
+from .model import textgenrnn_model
+import csv
+import re
+
+
+class textgenrnn:
+    META_TOKEN = '<s>'
+    config = {
+        'rnn_layers': 2,
+        'rnn_size': 128,
+        'rnn_bidirectional': False,
+        'max_length': 40,
+        'max_words': 10000,
+        'dim_embeddings': 100,
+        'word_level': False
+    }
+    default_config = config.copy()
+
+    def __init__(self, weights_path=None,
+                 vocab_path=None,
+                 config_path=None,
+                 name="textgenrnn"):
+
+        if weights_path is None:
+            weights_path = resource_filename(__name__,
+                                             'textgenrnn_weights.hdf5')
+
+        if vocab_path is None:
+            vocab_path = resource_filename(__name__,
+                                           'textgenrnn_vocab.json')
+
+        if config_path is not None:
+            with open(config_path, 'r',
+                      encoding='utf8', errors='ignore') as json_file:
+                self.config = json.load(json_file)
+
+        self.config.update({'name': name})
+        self.default_config.update({'name': name})
+
+        with open(vocab_path, 'r',
+                  encoding='utf8', errors='ignore') as json_file:
+            self.vocab = json.load(json_file)
+
+        self.tokenizer = Tokenizer(filters='', char_level=True)
+        self.tokenizer.word_index = self.vocab
+        self.num_classes = len(self.vocab) + 1
+        self.model = textgenrnn_model(self.num_classes,
+                                      cfg=self.config,
+                                      weights_path=weights_path)
+        self.indices_char = dict((self.vocab[c], c) for c in self.vocab)
+
+    def generate(self, n=1, return_as_list=False, prefix=None,
+                 temperature=0.5, max_gen_length=300):
+        gen_texts = []
+        for _ in range(n):
+            gen_text = textgenrnn_generate(self.model,
+                                           self.vocab,
+                                           self.indices_char,
+                                           prefix,
+                                           temperature,
+                                           self.config['max_length'],
[…]

### 2018-04-17 experiments/textgenrnn/textgenrnn/AttentionWeightedAverage.py
remove textgenrnn

@@ -1,65 +1,0 @@
-from keras.engine import InputSpec, Layer
-from keras import backend as K
-from keras import initializers
-import tensorflow as tf
-
-
-class AttentionWeightedAverage(Layer):
-    """
-    Computes a weighted average of the different channels across timesteps.
-    Uses 1 parameter pr. channel to compute the attention value for
-    a single timestep.
-    """
-
-    def __init__(self, return_attention=False, **kwargs):
-        self.init = initializers.get('uniform')
-        self.supports_masking = True
-        self.return_attention = return_attention
-        super(AttentionWeightedAverage, self).__init__(** kwargs)
-
-    def build(self, input_shape):
-        self.input_spec = [InputSpec(ndim=3)]
-        assert len(input_shape) == 3
-
-        self.W = self.add_weight(shape=(input_shape[2], 1),
-                                 name='{}_W'.format(self.name),
-                                 initializer=self.init)
-        self.trainable_weights = [self.W]
-        super(AttentionWeightedAverage, self).build(input_shape)
-
-    def call(self, x, mask=None):
-        # computes a probability distribution over the timesteps
-        # uses 'max trick' for numerical stability
-        # reshape is done to avoid issue with Tensorflow
-        # and 1-dimensional weights
-        logits = K.dot(x, self.W)
-        x_shape = K.shape(x)
-        logits = K.reshape(logits, (x_shape[0], x_shape[1]))
-        ai = K.exp(logits - K.max(logits, axis=-1, keepdims=True))
-
-        # masked timesteps have zero weight
-        if mask is not None:
-            mask = K.cast(mask, K.floatx())
-            ai = ai * mask
-        att_weights = ai / (K.sum(ai, axis=1, keepdims=True) + K.epsilon())
-        weighted_input = x * K.expand_dims(att_weights)
-        result = K.sum(weighted_input, axis=1)
-        if self.return_attention:
-            return [result, att_weights]
-        return result
-
-    def get_output_shape_for(self, input_shape):
-        return self.compute_output_shape(input_shape)
-
-    def compute_output_shape(self, input_shape):
-        output_len = input_shape[2]
-        if self.return_attention:
-            return [(input_shape[0], output_len), (input_shape[0],
-                                                   input_shape[1])]
-        return (input_shape[0], output_len)
-
-    def compute_mask(self, input, input_mask=None):
-        if isinstance(input_mask, list):
-            return [None] * len(input_mask)
-        else:
-            return None

### 2018-04-17 experiments/textgenrnn/textgenrnn/model.py

@@ -1,87 +1,0 @@
-from keras.optimizers import RMSprop
-from keras.layers import Input, Embedding, Dense, LSTM, Bidirectional
-from keras.layers import CuDNNLSTM, concatenate, Reshape
-from keras.models import Model
-from keras import backend as K
-from .AttentionWeightedAverage import AttentionWeightedAverage
-
-
-def textgenrnn_model(num_classes, cfg, context_size=None,
-                     weights_path=None,
-                     optimizer=RMSprop(lr=4e-3, rho=0.99)):
-    '''
-    Builds the model architecture for textgenrnn and
-    loads the specified weights for the model.
-    '''
-
-    input = Input(shape=(cfg['max_length'],), name='input')
-    embedded = Embedding(num_classes, cfg['dim_embeddings'],
-                         input_length=cfg['max_length'],
-                         name='embedding')(input)
-
-    rnn_layer_list = []
-    for i in range(cfg['rnn_layers']):
-        prev_layer = embedded if i is 0 else rnn_layer_list[-1]
-        rnn_layer_list.append(new_rnn(cfg, i+1)(prev_layer))
-
-    seq_concat = concatenate([embedded] + rnn_layer_list, name='rnn_concat')
-    attention = AttentionWeightedAverage(name='attention')(seq_concat)
-    output = Dense(num_classes, name='output', activation='softmax')(attention)
-
-    if context_size is None:
-        model = Model(inputs=[input], outputs=[output])
-        if weights_path is not None:
-            model.load_weights(weights_path, by_name=True)
-        model.compile(loss='categorical_crossentropy', optimizer=optimizer)
-
-    else:
-        context_input = Input(
-            shape=(context_size,), name='context_input')
-        context_reshape = Reshape((context_size,),
-                                  name='context_reshape')(context_input)
-        merged = concatenate([attention, context_reshape], name='concat')
-        main_output = Dense(num_classes, name='context_output',
-                            activation='softmax')(merged)
-
-        model = Model(inputs=[input, context_input],
-                      outputs=[main_output, output])
-        if weights_path is not None:
-            model.load_weights(weights_path, by_name=True)
-        model.compile(loss='categorical_crossentropy', optimizer=optimizer,
-                      loss_weights=[0.8, 0.2])
-
-    return model
-
-
-'''
-Create a new LSTM layer per parameters. Unfortunately,
-each combination of parameters must be hardcoded.
-
[…]

### 2018-04-17 experiments/textgenrnn/textgenrnn/textgenrnn.py

@@ -1,420 +1,0 @@
-from keras.callbacks import LearningRateScheduler, Callback
-from keras.models import Model, load_model
-from keras.preprocessing import sequence
-from keras.preprocessing.text import Tokenizer, text_to_word_sequence
-from keras import backend as K
-from sklearn.preprocessing import LabelBinarizer
-import numpy as np
-import json
-import h5py
-from pkg_resources import resource_filename
-from .model import textgenrnn_model
-import csv
-import re
-
-
-class textgenrnn:
-    META_TOKEN = '<s>'
-    config = {
-        'rnn_layers': 2,
-        'rnn_size': 128,
-        'rnn_bidirectional': False,
-        'max_length': 40,
-        'max_words': 10000,
-        'dim_embeddings': 100,
-        'word_level': False
-    }
-    default_config = config.copy()
-
-    def __init__(self, weights_path=None,
-                 vocab_path=None,
-                 config_path=None,
-                 name="textgenrnn"):
-
-        if weights_path is None:
-            weights_path = resource_filename(__name__,
-                                             'textgenrnn_weights.hdf5')
-
-        if vocab_path is None:
-            vocab_path = resource_filename(__name__,
-                                           'textgenrnn_vocab.json')
-
-        if config_path is not None:
-            with open(config_path, 'r',
-                      encoding='utf8', errors='ignore') as json_file:
-                self.config = json.load(json_file)
-
-        self.config.update({'name': name})
-        self.default_config.update({'name': name})
-
-        with open(vocab_path, 'r',
-                  encoding='utf8', errors='ignore') as json_file:
-            self.vocab = json.load(json_file)
-
-        self.tokenizer = Tokenizer(filters='', char_level=True)
-        self.tokenizer.word_index = self.vocab
-        self.num_classes = len(self.vocab) + 1
-        self.model = textgenrnn_model(self.num_classes,
-                                      cfg=self.config,
-                                      weights_path=weights_path)
-        self.indices_char = dict((self.vocab[c], c) for c in self.vocab)
-
-    def generate(self, n=1, return_as_list=False, prefix=None,
-                 temperature=0.5, max_gen_length=300):
-        gen_texts = []
-        for _ in range(n):
-            gen_text = textgenrnn_generate(self.model,
-                                           self.vocab,
-                                           self.indices_char,
-                                           prefix,
-                                           temperature,
-                                           self.config['max_length'],
[…]

### 2018-04-17 experiments/dataset.py
attempt textgenrnn archetecture

@@ -33,7 +33,7 @@ def remove_mentions(vTEXT):
     return vTEXT
 
 
-ALPHABET = frozenset(string.ascii_lowercase + string.punctuation + ' #@')
+ALPHABET = frozenset(string.ascii_letters + string.punctuation + ' #@')
 
 
 def preprocess(text, alphabet):
@@ -63,6 +63,8 @@ def load_dataset(files, verbose=True):
     dataset = (t['text']for t in dataset)
     # Remove URLs from the tweets.
     dataset = (remove_urls(t) for t in dataset)
+    # # Remove @mentions from tweets.
+    # dataset = (remove_mentions(t) for t in dataset)
     # Preprocess each tweet, filtering out nonascii alphabetic
     dataset = (''.join(preprocess(t, ALPHABET)) for t in dataset)
 

### 2018-04-17 experiments/model.py

@@ -4,8 +4,10 @@ used by the rest of the workflow.
 """
 
 from datetime import datetime
-from keras.layers import LSTM, Dense, BatchNormalization
+# from keras.layers import LSTM, Dense, BatchNormalization
 from keras.models import Sequential, model_from_json
+
+from textgenrnn.textgenrnn import textgenrnn
 
 # Turn the corpus into sequences 60 characters long
 SEQ_LEN = 60
@@ -25,28 +27,29 @@ EPOCHS = 50
 BASENAME = 'models/64x6-rms-batchnorm-50'
 
 
-def build_model(seq_length, num_chars, verbose):
+def build_model(verbose):
     """
         Builds an LSTM model for generating text.
     """
     if verbose:
         print('Building model...')
-    model = Sequential()
-    model.add(LSTM(units=64, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
-    model.add(BatchNormalization())
-    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    model.add(BatchNormalization())
-    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    model.add(BatchNormalization())
-    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    model.add(BatchNormalization())
-    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    model.add(BatchNormalization())
-    model.add(LSTM(units=64, unit_forget_bias=True))
-    model.add(BatchNormalization())
-    model.add(Dense(num_chars, activation='softmax'))
+    # model = Sequential()
+    # model.add(LSTM(units=64, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
+    # model.add(BatchNormalization())
+    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    # model.add(BatchNormalization())
+    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    # model.add(BatchNormalization())
+    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    # model.add(BatchNormalization())
+    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    # model.add(BatchNormalization())
+    # model.add(LSTM(units=64, unit_forget_bias=True))
+    # model.add(BatchNormalization())
[…]

### 2018-04-17 experiments/train.py

@@ -74,32 +74,48 @@ def plot_model_loss(filename, hist, verbose=True):
 
 def main(verbose):
     dataset = load_dataset(glob('../data/trump_tweet_data_archive/condensed_*.json.zip'), verbose)
-    corpus, sequences, next_chars, c2i, i2c, nc = seq_data(dataset, SEQ_LEN, SEQ_STEP, verbose)
+    # corpus, sequences, next_chars, c2i, i2c, nc = seq_data(dataset, SEQ_LEN, SEQ_STEP, verbose)
 
-    if verbose:
-        print(f'corpus length: {len(corpus)}')
-        print(f'num characters: {nc}')
-        print(f'number of sequences: {len(sequences)}')
+    # if verbose:
+    #     print(f'corpus length: {len(corpus)}')
+    #     print(f'num characters: {nc}')
+    #     print(f'number of sequences: {len(sequences)}')
 
-    # The data is shuffled so the validation data isn't simply the latest 20% of tweets
-    X, y = vec_data(sequences, next_chars, SEQ_LEN, nc, c2i, verbose)
-    n = len(X)
-    num_val = int(PERCENT_VALIDATION * n)
-    X_val = X[n - num_val:]
-    y_val = y[n - num_val:]
+    # # The data is shuffled so the validation data isn't simply the latest 20% of tweets
+    # X, y = vec_data(sequences, next_chars, SEQ_LEN, nc, c2i, verbose)
+    # n = len(X)
+    # num_val = int(PERCENT_VALIDATION * n)
+    # X_val = X[n - num_val:]
+    # y_val = y[n - num_val:]
 
-    X_train = X[:n - num_val]
-    y_train = y[:n - num_val]
+    # X_train = X[:n - num_val]
+    # y_train = y[:n - num_val]
 
-    if verbose:
-        print(f'Number validation samples: {num_val}')
+    # if verbose:
+    #     print(f'Number validation samples: {num_val}')
 
-    model = build_model(SEQ_LEN, nc, verbose)
-    history = train_model(model, X_train, y_train, X_val, y_val, verbose)
-    plot_model_loss(BASENAME, history, verbose)
-    save_model(model, BASENAME, verbose)
-    # Generate sample tweets using 10 random seeds from the corpus.
-    generate(BASENAME, model, corpus, c2i, i2c, nc, 10, verbose)
+    # model = build_model(SEQ_LEN, nc, verbose)
[…]

### 2018-04-17 experiments/dataset.py
fix training data structure

@@ -40,7 +40,7 @@ def preprocess(text, alphabet):
     """
         Preprocess text using the frozenset `alphabet` given.
     """
-    return filter(alphabet.__contains__, text.lower())
+    return filter(alphabet.__contains__, text)
 
 def load_dataset(files, verbose=True):
     """
@@ -66,7 +66,7 @@ def load_dataset(files, verbose=True):
     # # Remove @mentions from tweets.
     # dataset = (remove_mentions(t) for t in dataset)
     # Preprocess each tweet, filtering out nonascii alphabetic
-    dataset = (''.join(preprocess(t, ALPHABET)) for t in dataset)
+    dataset = [''.join(preprocess(t, ALPHABET)) for t in dataset]
 
     return dataset
 

### 2018-04-17 experiments/train.py

@@ -104,7 +104,8 @@ def main(verbose):
     model = build_model(verbose=True)
     model.train_new_model(
         dataset,
-        context_labels=[],
+        # Label each tweet as a Trump tweet. Useful when combining multiple sources
+        context_labels=['trump' for t in dataset],
         num_epochs=5,
         gen_epochs=1,
         batch_size=128,

### 2018-04-18 experiments/train.py
tweak training script

@@ -25,6 +25,8 @@ from generate import generate
 from model import (BASENAME, BATCH_SIZE, EPOCHS, PERCENT_VALIDATION,
                    SEQ_LEN, SEQ_STEP, build_model, save_model)
 
+from textgenrnn.textgenrnn import textgenrnn
+
 # Hack to keep keras from allocating the whole damn gpu.
 config = tf.ConfigProto()
 config.gpu_options.allow_growth = True
@@ -101,12 +103,12 @@ def main(verbose):
     # # Generate sample tweets using 10 random seeds from the corpus.
     # generate(BASENAME, model, corpus, c2i, i2c, nc, 10, verbose)
 
-    model = build_model(verbose=True)
-    model.train_new_model(
+    gen = textgenrnn()
+    gen.train_new_model(
         dataset,
         # Label each tweet as a Trump tweet. Useful when combining multiple sources
         context_labels=['trump' for t in dataset],
-        num_epochs=5,
+        num_epochs=6,
         gen_epochs=1,
         batch_size=128,
         prop_keep=1.0,
@@ -117,6 +119,7 @@ def main(verbose):
         dim_embeddings=100,
         word_level=False,
     )
+    gen.generate(10, temperature=0.2)
 
 
 if __name__ == '__main__':

### 2018-04-19 paper/paper.tex
Add TODO notes to paper

@@
 
-\title{Generating haikus with an LSTM network}
+\title{Generating presidential tweets with an LSTM network}
 
-\author{
+\author{%
     Austin Gill\\
@@
     Rapid City SD, 57701\\
-    \texttt{\href{mailto:austin.gill@mines.sdsmt.edu}{austin.gill@mines.sdsmt.edu}} \\
+    \texttt{\href{mailto:austin.gill@mines.sdsmt.edu}{austin.gill@mines.sdsmt.edu}}\\
 }
@@
 \begin{abstract}
-    Here's an abstract.
+    \todoinline{Write an abstract, but after the paper is finished.}
 \end{abstract}
@@
 \section{Introduction}
-    Here's the introduction.
+    \todoinline{
+        \begin{itemize}
+            \item Purpose
+            \item Background and general research (Haiku paper using word association norms?)
+            \item Mention original intent of generating haikus?
+            \item Introduce and defend the final model
+            \item Introduce the experimental method used to verify the model
+        \end{itemize}
+    }
 
-\section{Stuff}
-    Here's some stuff you didn't know!
+\section{Experiments}
+    \todoinline{
+        \begin{itemize}
+            \item Initial attempt based on generation of a large book (inspired by\dots)
+            \item Formulation of the input/output/generation data form? (How to sample with some diversity)
+            \item Different architectures attempted (Different plots, but the plots are all pngs\dots)
+            \item Final architecture inspired by some paper that I have to find
+            \item number epochs, optimizer, etc.
+        \end{itemize}
+    }
 
-\section{Conclusion}
-    Conclude something\todo{Something useful!}
+\section{Results}
+    \todoinline{
+        \begin{itemize}
+            \item Hack validation data into training and produce plots?
+            \item Give example tweets?
+            \item Grammar, spelling, etc.
+            \item Accuracy is meaningless because you're trying to create \textit{new} tweets.
+This is where the sampling with diversities comes in.
+The higher diversity the more creative the network is, but the more meaningless it becomes.
+            \item Produce a largeish amount of generated tweets.
+        \end{itemize}
+    }
 
+\section{Discussion}
+    \todoinline{
+        \begin{itemize}
+            \item How good was it?
+            \item Accuracy meaningless because attempt is creative generation.
+            \item Did it pick up on patterns backed by data?
+        \end{itemize}
+    }
+
+\section{Future Work}
+    \todoinline{
+        \begin{itemize}
+            \item Word association norms?
+            \item Embeddings?
+            \item Combining sources?
+            \item Grammar and spelling
+            \item GAN
+            \item Crowd-sourced adversarial network as a training component?
+        \end{itemize}
+    }
+
+\bibliographystyle{ieeetr}
+% Automatically places a \section*{References}...
+\bibliography{paper}
 \end{document}

### 2018-04-21 paper/paper.tex
Add Introduction and Experiments section.
wrapfig is a whole new kind of hell that I wouldn't wish on anyone, even
on my ex-girlfriend.

@@
 \section{Introduction}
-    \todoinline{
-        \begin{itemize}
-            \item Purpose
-            \item Background and general research (Haiku paper using word association norms?)
-            \item Mention original intent of generating haikus?
-            \item Introduce and defend the final model
-            \item Introduce the experimental method used to verify the model
-        \end{itemize}
-    }
+    The field of classifying text with machine learning models is well-studied.
+However, the field of text-generational models is still being actively researched.
+Generational models are interesting because there is no true idea of ``accuracy'' because the models are \textit{creative}.
+Given some input, a generational model does not seek to produce an output identical to that of the training set.
+Instead, there is a subtle tug-of-war between reproducing text samples found in the training set and producing \textit{new} text samples that sound similar.
+
+    The author's original goal was to produce a machine learning model to generate haikus, a short form of Japanese poetry.
+\citet{yael_2010} used word association norms to extend previous generational models to generate haikus with some success.
+However, the author was unable to scrape together a large enough data set to build a model.
+Additionally, \citet{yael_2010} did not specify their model architecture in detail, leaving more research to be done than the author had time for.
+
+    However, there have been a number of open source implementations of generational models for larger pieces of text.
+\citet{thoutt_2017} built a generational model to generate the sixth \textit{A Song of Ice and Fire} \citep{martin} book due to its untimely publication.
+
+    Finding an interesting and relevant text generation problem with a large corpus to train a model on was as easy as turning to Twitter.
+There has been recent research using Twitter.
+\citet{felbo_2017} produced a model to perform sentiment, emotion, and sarcasm analysis on tweets taking into account their use of Emoji characters.
+\citet{woolf_2017} constructed a sophisticated model to generate similar-sounding tweets.
+However, \citet{woolf_2017} also attempted (with success) to train on the tweets of multiple Twitter users, and used contextual tagging to generate tweets combining the semantics of multiple (and possibly quite differently behaving) users.
+Twitter provides a fairly simple API to download a user's tweets.
+However, the API limits the number tweets available to the 3200 most recent tweets.
+This was a smaller data set than the haikus the author was able to scrape together.
+
+    Luckily we live in a time where passionate people can put an extraordinary amount of effort in building a publicly accessible archive of tweets by prominent politicians.
+\citet{brown_2017} produced a public archive of President Donald J.
+Trump's tweets and hosted it on GitHub for public use.
+The archive is updated hourly, and does to remove deleted tweets.
+
+    Before deep learning's rise to fame, text generation was done using probabilistic Markov models where the next word in a sequence was generated based on the current (and only the current) state in the Markov chain.
+These models lacked enough context to construct cogent text samples reliably.
+Recurrent machine learning models are the current response to providing contextual information.
+
+    Rather than generate the next token in a sequence based only on the previous token, recurrent models provide some amount of history so that the training can pick up on larger patterns in the training data, typically referred to as the corpus when dealing with text.
+\citet{karpathy_2015} has done quite a bit of research using recurrent models with a generational purpose.
+For example, \citet{karpathy_2015} constructed recurrent models trained on the Linux kernel source code, as well as various mathematical research papers, to produce various C and \LaTeX{} code that almost compiled!
+
+    The current standard model in text generational models, according to \citet{karpathy_2015}, is some number of Long Short-Term Memory (LSTM) layers followed by a single densely connected output layer with a softmax activation function to provide a one-hot token encoding vector as an output.
+
+    As mentioned though, accuracy does not make a useful metric to monitor training progression because generational models do not seek to reproduce the training set.
+Instead, they attempt to produce a new corpus that has some amount of similarity.
+In scope, the author did not have the resources to provide an in-depth similarity analysis, and instead relied on a subjective personal measure of how well the produced models performed.
 
 \section{Experiments}
-    \todoinline{
-        \begin{itemize}
-            \item Initial attempt based on generation of a large book (inspired by\dots)
-            \item Formulation of the input/output/generation data form? (How to sample with some diversity)
-            \item Different architectures attempted (Different plots, but the plots are all pngs\dots)
-            \item Final architecture inspired by some paper that I have to find
-            \item number epochs, optimizer, etc.
-        \end{itemize}
-    }
+    \subsection{Inspiration}
+        Most heavily, the author was inspired by previous work generating large corpuses by \citet{thoutt_2017} and the work on recurrent architectures by \citet{karpathy_2015}.
+In short, the author chose to deal almost exclusively with LSTM layers in some configuration with a one-hot encoded softmax output.
+
+        Due to ease-of-use and a simple high-level interface, the author chose to implement all models using Keras \citep{chollet2015keras}.
+Keras provided an acceptably abstracted view of the models that enabled to author to focus on the model architecture rather than each model's implementation.
+
+    \subsection{Data Formatting}
+        After retrieving the tweets from \citet{brown_2017}'s archive, the next difficulty was deciding how to effectively format the data to facilitate its use in a neural network.
+Based on the Keras documentation and LSTM generational examples, the author decided to follow the following three steps
+
+        \begin{enumerate}
+            \item Preprocess the data to remove Emojis, retweets, capitalization, URLs, images, etc.
+
+            Other tweet generation projects remove username ``\texttt{@mentions}'' from their training set, but the author felt it was valuable to not strip them out.
+
+            \item Concatenate each tweet to form a larger corpus to sample from.
+            \item Chunk the corpus into overlapping sequences of tokens.
+This forms the inputs to the network.
+The expected output is the token immediately following the end of each sequence.
+Each sequence was 60 tokens long and separated by 3 tokens as shown in~\autoref{fig:sequences}
+            \item Vectorize each sequence by mapping each token to a vector with a \texttt{TRUE} value in the index corresponding to that token.
+        \end{enumerate}
+
+        \begin{figure}[h]
+            \centering
+            \begin{tikzpicture}[scale=0.25]
+
+                \draw[thick] (0, 0) rectangle (40, 3);
+                \draw[thick] (7, 0) rectangle (47, 3);
+                \draw [decorate,
+                       thick,
+                       decoration={brace, amplitude=6pt, mirror},
+                       xshift=0pt,
+                       yshift=-10pt]
+                    (0, 0) -- node[midway, below, yshift=-5] {3 tokens} (6.5, 0);
+                \draw [decorate,
+                       thick,
+                       decoration={brace, amplitude=6pt},
+                       xshift=0pt,
+                       yshift=10pt]
+                    (0, 3) -- node[midway, above, yshift=5] {Sequence $n$, 60 tokens} (40, 3);
+                \draw [decorate,
+                       thick,
+                       decoration={brace, amplitude=6pt, mirror},
+                       xshift=0pt,
+                       yshift=-10pt]
+                 (7.5, 0) -- node[midway, below, yshift=-5] {Sequence $n + 1$, 60 tokens} (47, 0);
+            \end{tikzpicture}
+            \caption{Chunking the corpus into overlapping sequences}\label{fig:sequences}
+        \end{figure}
+
+        There are two options for tokenizing the corpus.
+The author chose to tokenize by character rather than by word.
+This means that the vectorized sequences are sequences of characters, and the model is predicting the very next character of each sequence.
+
+        \newpage
+        % I have a new hatred for floating wrapped figures. Welcome to a new kind of hell.
+        \begin{wrapfigure}[50]{r}[20pt]{0pt}
+            \centering
+            \includegraphics[width=0.39\textwidth]{figures/64x6.eps}
+            \caption{A deep and narrow LSTM model using batch normalization}\label{fig:archetecture-64x6}
+        \end{wrapfigure}
+
+        Additionally, the sequences overlap a great deal, and contain an enormous amount of redundant data.
+The dataset in JSON form takes a mere 9.3 megabytes, but in vectorized form takes almost 5 gigabytes.
+Also note that even though there are only $\sim 37$ thousand tweets by President Trump, the vectorized dataset turns into $\sim 1.1$ \textit{million} sequences.
+
+        Also note that the author trained the experimental models on \textit{all} of President Trump's tweets from his account creation in 2009 to 2018.\todo{Move this to discussion}{}
+
+    \subsection{Model Architectures}
+        The author attempted 21 different variations of LSTM models in both height and depth.
+Most examples and online recommendations were to use relatively shallow networks with height determined by circumstance.\todo{Citation needed?}{}
+
+        One such configuration that showed promise is shown in~\autoref{fig:archetecture-64x6}.
+This model was composed of six 64-unit LSTM layers with batch normalization.
+Each configuration used 60 character long sequences as the model input, and a one-hot encoded vector output to indicate the most likely next character in the sequence.
+
+        While the author experimented with regularization, overfitting was not an issue due to the sheer size of the training set.
+Thus, most of the best performing models used batch normalization rather than some form of dropout.
+
+        Note that the author was not able to significantly outperform his first attempt, which consisted of a single 256-unit LSTM layer without batch normalization or dropout.
+
+    \subsection{Training Variations}
+        The author initially used the Root Mean Square Propagation (RMSProp) optimization algorithm on recommendation, but also experimented with Stochastic Gradient Decent (SGD) optimizer.
+RMSProp is an adaptive optimizer that tunes its learning rate as it runs.
+The author favors RMSProp due to a considerably faster training time in both time elapsed and number of epochs necessary when compared with SGD.
+
+        When using RMSProp, it was typically only necessary to train for around 5 epochs, depending on the model size, while SGD (even with a higher learning rate) took on average 100 epochs to get comparable results.
+
+    \subsection{Generating New Tweets}
+        The neural network takes in a sequence of the training set, and predicts the next token in the corpus.
+To take a trained network and generate new tweets, we first begin with a seed randomly taken from the corpus.
+The network predicts the next token in the sequence as an array of probabilities, which is then sampled with some diversity because the intent of the model is to generate new creative work not found in the training set.
+
+        The newly generated token is appended to the seed, which is then shifted down one token.
+The author chose to repeat this process until the generated portion of the text (excluding the seed) is 280 characters long.
+
+        This process is repeated using several different probability sampling diversities for each randomly chosen seed, resulting in a different piece of generated text for each diversity.
+The higher the diversity, the more creative the model became, but at the cost of becoming less and less intelligible.
+The lower the diversity, the more intelligible became, but at the cost of simply repeating portions of text appearing in the corpus.
 
@@
     }
+    \subsection{256 Unit, 1 Layer LSTM}
+
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.4\textwidth]{figures/256dropout.eps}
+            \caption{256 unit LSTM model with dropout}\label{fig:archetecture-256}
+        \end{figure}
+
+    \subsection{64 Unit, 6 Layer LSTM}
+        The following results are using the model shown in~\autoref{fig:archetecture-64x6}.
+
+    \subsection{1024 Unit, 1 Layer LSTM}
+    \begin{figure}[h]
+        \centering
+        \includegraphics[width=0.5\textwidth]{figures/1024batchnorm.eps}
+        \caption{1024 unit LSTM model with batch normalization}\label{fig:archetecture-256}
+    \end{figure}
+
+    \subsection{Textgenrnn}
+    \begin{figure}[h]
+        \centering
+        \includegraphics[width=\textwidth]{figures/context_model.png}
+        \caption{Textgenrnn architecture}\label{fig:archetecture-256}
+    \end{figure}
 
@@
         \begin{itemize}
+            \item Word level generation. Instead of each character being mapped to an integer, create a vocab list and map each word to an integer.
             \item Word association norms?
@@
             \item Combining sources?
-            \item Grammar and spelling
-            \item GAN
+            \item Grammar and spelling, part of speech tagging, word association, linguistic features as a component of a GAN
             \item Crowd-sourced adversarial network as a training component?
@@
 
-\bibliographystyle{ieeetr}
+% Have to use one of the natbib styles. They all suck...
+\bibliographystyle{abbrvnat}
 % Automatically places a \section*{References}...

### 2018-04-22 experiments/dataset.py
Reset scripts to my model

@@ -40,7 +40,7 @@ def preprocess(text, alphabet):
     """
         Preprocess text using the frozenset `alphabet` given.
     """
-    return filter(alphabet.__contains__, text)
+    return filter(alphabet.__contains__, text.lower())
 
 def load_dataset(files, verbose=True):
     """

### 2018-04-22 experiments/model.py

@@ -4,10 +4,13 @@ used by the rest of the workflow.
 """
 
 from datetime import datetime
-# from keras.layers import LSTM, Dense, BatchNormalization
+from keras.layers import LSTM, Dense, BatchNormalization
+# A CUDA accelerated LSTM layer to *really* improve training time. Cannot
+# Blindly swap out, needs to be retrained :(
+# from keras.layers import CuDNNLSTM as LSTM
 from keras.models import Sequential, model_from_json
 
-from textgenrnn.textgenrnn import textgenrnn
+# from textgenrnn.textgenrnn import textgenrnn
 
 # Turn the corpus into sequences 60 characters long
 SEQ_LEN = 60
@@ -27,29 +30,29 @@ EPOCHS = 50
 BASENAME = 'models/64x6-rms-batchnorm-50'
 
 
-def build_model(verbose):
+def build_model(seq_length, num_chars, verbose):
     """
         Builds an LSTM model for generating text.
     """
     if verbose:
         print('Building model...')
-    # model = Sequential()
-    # model.add(LSTM(units=64, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
-    # model.add(BatchNormalization())
-    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    # model.add(BatchNormalization())
-    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    # model.add(BatchNormalization())
-    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    # model.add(BatchNormalization())
-    # model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
-    # model.add(BatchNormalization())
-    # model.add(LSTM(units=64, unit_forget_bias=True))
-    # model.add(BatchNormalization())
-    # model.add(Dense(num_chars, activation='softmax'))
+    model = Sequential()
+    model.add(LSTM(units=64, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
+    model.add(BatchNormalization())
+    model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
[…]

### 2018-04-22 experiments/train.py

@@ -25,7 +25,7 @@ from generate import generate
 from model import (BASENAME, BATCH_SIZE, EPOCHS, PERCENT_VALIDATION,
                    SEQ_LEN, SEQ_STEP, build_model, save_model)
 
-from textgenrnn.textgenrnn import textgenrnn
+# from textgenrnn.textgenrnn import textgenrnn
 
 # Hack to keep keras from allocating the whole damn gpu.
 config = tf.ConfigProto()
@@ -76,50 +76,50 @@ def plot_model_loss(filename, hist, verbose=True):
 
 def main(verbose):
     dataset = load_dataset(glob('../data/trump_tweet_data_archive/condensed_*.json.zip'), verbose)
-    # corpus, sequences, next_chars, c2i, i2c, nc = seq_data(dataset, SEQ_LEN, SEQ_STEP, verbose)
+    corpus, sequences, next_chars, c2i, i2c, nc = seq_data(dataset, SEQ_LEN, SEQ_STEP, verbose)
 
-    # if verbose:
-    #     print(f'corpus length: {len(corpus)}')
-    #     print(f'num characters: {nc}')
-    #     print(f'number of sequences: {len(sequences)}')
+    if verbose:
+        print(f'corpus length: {len(corpus)}')
+        print(f'num characters: {nc}')
+        print(f'number of sequences: {len(sequences)}')
 
-    # # The data is shuffled so the validation data isn't simply the latest 20% of tweets
-    # X, y = vec_data(sequences, next_chars, SEQ_LEN, nc, c2i, verbose)
-    # n = len(X)
-    # num_val = int(PERCENT_VALIDATION * n)
-    # X_val = X[n - num_val:]
-    # y_val = y[n - num_val:]
+    # The data is shuffled so the validation data isn't simply the latest 20% of tweets
+    X, y = vec_data(sequences, next_chars, SEQ_LEN, nc, c2i, verbose)
+    n = len(X)
+    num_val = int(PERCENT_VALIDATION * n)
+    X_val = X[n - num_val:]
+    y_val = y[n - num_val:]
 
-    # X_train = X[:n - num_val]
-    # y_train = y[:n - num_val]
+    X_train = X[:n - num_val]
+    y_train = y[:n - num_val]
 
-    # if verbose:
-    #     print(f'Number validation samples: {num_val}')
+    if verbose:
+        print(f'Number validation samples: {num_val}')
 
-    # model = build_model(SEQ_LEN, nc, verbose)
[…]

### 2018-04-22 paper/paper.tex
Add finished paper.
I've never hated a paper before. This is a new feeling.

@@
 \begin{abstract}
-    \todoinline{Write an abstract, but after the paper is finished.}
+    We motivate and develop a recurrent generational model to generate tweets similar to those of President Donald J.
+Trump.
+We outline previous research in generational models, and point out the difficulties of objective measurement of inherently subjective creative models.
+After experimenting with several recurrent models, we settle on a 6 layer, 64 unit LSTM model, and recommend a more advanced model as proposed by other research.
+We finally recommend using advanced natural language processing techniques to develop a Generative Adversarial Network in any future work.
 \end{abstract}
@@
 
-        \begin{enumerate}
-            \item Preprocess the data to remove Emojis, retweets, capitalization, URLs, images, etc.
-
-            Other tweet generation projects remove username ``\texttt{@mentions}'' from their training set, but the author felt it was valuable to not strip them out.
-
-            \item Concatenate each tweet to form a larger corpus to sample from.
-            \item Chunk the corpus into overlapping sequences of tokens.
-This forms the inputs to the network.
-The expected output is the token immediately following the end of each sequence.
-Each sequence was 60 tokens long and separated by 3 tokens as shown in~\autoref{fig:sequences}
-            \item Vectorize each sequence by mapping each token to a vector with a \texttt{TRUE} value in the index corresponding to that token.
-        \end{enumerate}
-
-        \begin{figure}[h]
+        \begin{figure}[H]
             \centering
@@
 
+        \begin{enumerate}
+            \item Preprocess the data to remove Emojis, retweets, capitalization, URLs, images, etc.
+
+            Other tweet generation projects remove username ``\texttt{@mentions}'' from their training set, but the author felt it was valuable to not strip them out.
+
+            \item Concatenate each tweet to form a larger corpus to sample from.
+            \item Chunk the corpus into overlapping sequences of tokens.
+This forms the inputs to the network.
+The expected output is the token immediately following the end of each sequence.
+Each sequence was 60 tokens long and separated by 3 tokens as shown in~\autoref{fig:sequences}
+            \item Vectorize each sequence by mapping each token to a vector with a \texttt{TRUE} value in the index corresponding to that token.
+        \end{enumerate}
+
+        % I have a new hatred for floating wrapped figures. Welcome to a new kind of hell.
+        % Also, tall and skinny figures just aren't a good idea...
+        \begin{wrapfigure}[47]{r}[40pt]{0pt}
+            \centering
+            \includegraphics[width=0.36\textwidth]{figures/64x6.eps}
+            \caption{A deep and narrow LSTM model using batch normalization}\label{fig:architecture-64x6}
+        \end{wrapfigure}
+
         There are two options for tokenizing the corpus.
@@
 This means that the vectorized sequences are sequences of characters, and the model is predicting the very next character of each sequence.
-
-        \newpage
-        % I have a new hatred for floating wrapped figures. Welcome to a new kind of hell.
-        \begin{wrapfigure}[50]{r}[20pt]{0pt}
-            \centering
-            \includegraphics[width=0.39\textwidth]{figures/64x6.eps}
-            \caption{A deep and narrow LSTM model using batch normalization}\label{fig:archetecture-64x6}
-        \end{wrapfigure}
 
@@
 
-        Also note that the author trained the experimental models on \textit{all} of President Trump's tweets from his account creation in 2009 to 2018.\todo{Move this to discussion}{}
+        Also note that the author trained the experimental models on \textit{all} of President Trump's tweets from his account creation in 2009 to 2018.
 
@@
         The author attempted 21 different variations of LSTM models in both height and depth.
-Most examples and online recommendations were to use relatively shallow networks with height determined by circumstance.\todo{Citation needed?}{}
+Most examples and online recommendations were to use relatively shallow networks with height determined by circumstance.
 
-        One such configuration that showed promise is shown in~\autoref{fig:archetecture-64x6}.
+        One such configuration that showed promise is shown in~\autoref{fig:architecture-64x6}.
 This model was composed of six 64-unit LSTM layers with batch normalization.
@@
         When using RMSProp, it was typically only necessary to train for around 5 epochs, depending on the model size, while SGD (even with a higher learning rate) took on average 100 epochs to get comparable results.
+All attempts used a categorical cross-entropy loss function due to the network output being one-hot encoded.
 
@@
 \section{Results}
-    \todoinline{
-        \begin{itemize}
-            \item Hack validation data into training and produce plots?
-            \item Give example tweets?
-            \item Grammar, spelling, etc.
-            \item Accuracy is meaningless because you're trying to create \textit{new} tweets.
-This is where the sampling with diversities comes in.
-The higher diversity the more creative the network is, but the more meaningless it becomes.
-            \item Produce a largeish amount of generated tweets.
-        \end{itemize}
-    }
+    The results for a few select architectures are summarized below.
+
     \subsection{256 Unit, 1 Layer LSTM}
+        \autoref{fig:256-loss} shows the training loss over time for a basic one-layer 256 unit LSTM model.
+The author used validation data to get a sense of how well the models generalized.
+The author did \textit{not} use a test set to evaluate how the model performed at the end of training because of the problem nature.
+Generational models are not intended to accurately map the training inputs to the training outputs.
+Thus evaluating the model accuracy on an unseen dataset gives no useful information.
 
@@
             \centering
-            \includegraphics[width=0.4\textwidth]{figures/256dropout.eps}
-            \caption{256 unit LSTM model with dropout}\label{fig:archetecture-256}
+            \includegraphics[width=0.7\textwidth]{figures/256dropout20.png}
+            \caption{Basic 256 unit LSTM model}\label{fig:256-loss}
         \end{figure}
 
+        Instead, the author used subjective measures to evaluate whether a model performed well or not.
+The output of this network, for lower sampling diversities, was quite intelligible.
+Several generated tweets have been cherry-picked, as \citet{karpathy_2015} notes is necessary for recurrent generational models, and shown below.
+
+        Note that all capitalization has been done by hand to make the results more legible.
+
+        \begin{quote}
+            ``The debate to the worst and the worst President Obama was a great president of the world is a great person. The Republicans will be a great job on the world''
+
+            ``The state of the state of the world is a great person. The Republicans will be on the world is a great poll to be the best interview discussing the Trump Tower''
+
+            ``@realDonaldTrump is the only one of the worst president of the world is a great person. The Republicans will be a great president of the world''
+
+            ``I have a great people of the worst and the debate to the worst President Obama is a great person.
+The Democrats are all talk about me to be the best interview discussing the Trump Tower''
+
+            ``The poor for their presidential party of his presidency to start doing for millions of the conservative and brand who has got a state of the good country!
+We need a state who says Hillary she are an amazing''
+        \end{quote}
+
+        It is quite evident that the network is catching on to certain speech patterns, but is completely blind to grammar, and only strings together cogent thoughts over small stretches of the generated text.
+When considering the sentence structure as a whole, it lacks meaning, but certain substrings (e.g., ``the worst President Obama was a great president'', or ``The Democrats are all talk about me'') form coherent thoughts.
+
     \subsection{64 Unit, 6 Layer LSTM}
-        The following results are using the model shown in~\autoref{fig:archetecture-64x6}.
+        The following results are using the model shown in \autoref{fig:architecture-64x6}.
+Note that \autoref{fig:64x6-loss} shows that this model took more training time to reach the same behavior as shown in \autoref{fig:256-loss}.
+This is expected because it's a larger model.
+
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.7\textwidth]{figures/64x6-rms-batchnorm-50.png}
+            \caption{64 unit, 6 layer LSTM model with batch normalization}\label{fig:64x6-loss}
+        \end{figure}
+
+        Again, several cherry-picked tweets are given below with capitalization added.
+
+        \begin{quote}
+            ``@MittRomney should be the only one who need it. Good move for the Scott exciting to Russia on @FoxNews for the administration policies commissioner''
+
+            ``Dems is a total loser with the Trump country is a great guy to the great states of The Apprentice''
+
+            ``@BarackObama scandaled will be Donald Trump liars will be a game on @foxandfriends at Americans and winner!''
+
+            ``Bad interview to the Trump country is the states of the first story and states of the Trump International Hotel in the country will be interviewed on @FoxNews and never win in the presidential''
+
+            ``Statements are something to be a truly great guy to the only one who can speak at the Trump Tower to see the president of the US''
+
+            ``Trump country is a great president of the USA and the art of the Democrats are something to be a great support the Trump Tower to make America great again!''
+
+            ``The end of Monday at your cards classing a billionaire inspiration''
+
+            ``And the Trump country is a great president of the USA and the Trump signing to stop president of the USA and the Trump signing to stop and the Trump signitictor is a great president of the USA and states and states of the Trump country is a great president of the USA''
+        \end{quote}
 
     \subsection{1024 Unit, 1 Layer LSTM}
+        During the experimentation process, it seemed like the shallow layers produced the most intelligible results, so the author experimented by making the models much much wider.
+However, the results were not promising.
+\autoref{fig:1024-loss} shows the training loss over time, which shows that the optimizer struggled with such a large model.
+Further, training even for 20 epochs took almost 24 hours on an NVIDA GTX 1080.
+
         \begin{figure}[h]
             \centering
-        \includegraphics[width=0.5\textwidth]{figures/1024batchnorm.eps}
-        \caption{1024 unit LSTM model with batch normalization}\label{fig:archetecture-256}
+            \includegraphics[width=0.7\textwidth]{figures/1024-rms-batchnorm-20.png}
+            \caption{1024 unit LSTM model with batch normalization}\label{fig:1024-loss}
         \end{figure}
 
+        The generated results are given below.
+
+        \begin{quote}
+            ``and and and and and and and the and and and and and the and and the and and and and and and and and and and and and and the and and and the and the and the and and the and and the and and and the and and and and and and the and the and than the and and the and and and and and''
+
+            ``@realDonaldTrump i we aa seare to need a fight and and any every is a tered beanies and the of the real in the life and that I live a great and leader I love to the beal about is any and and the me.
+The thanes and that I and and and lime''
+
+            ``America get I and the thang is and the ane that I at see any I and than Obama in is and peolle last is than the antary this peesident a.
+The unition so to thanised is a taine in the really and it in I've any way a deal to the for the aldonal and was and we will be a like''
+        \end{quote}
+
     \subsection{Textgenrnn}
+        After it was too late to incorporate any meaningful changes, the author was directed to a GitHub project \citep{woolf_2017} built expressly for generating text using a recurrent network.
+The model architecture is shown in \autoref{fig:architecture-textgenrnn}.
+This architecture is interesting because it has a context input to allow mixing tagging different pieces of the training set with context.
+The example usage given in the documentation indicates that this works well when combining text written by different authors.
+
         \begin{figure}[h]
@@
             \includegraphics[width=\textwidth]{figures/context_model.png}
-        \caption{Textgenrnn architecture}\label{fig:archetecture-256}
+            \caption{Textgenrnn architecture}\label{fig:architecture-textgenrnn}
         \end{figure}
 
+        For completeness, Textgenrnn, as trained by the author, may be used as follows
+
+        \begin{minted}{python}
+            from textgenrnn.textgenrnn import textgenrnn  # Installable with Pip
+            # Files found in trumperator/experiments/
+            t = textgenrnn(config_path='trumperator_config.json',
+                           vocab_path='trumperator_vocab.json',
+                           weights_path='trumperator_weights.hdf5')
+            t.generate(1, temperature=0.2)
+        \end{minted}
+
+        Textgenrnn has the option of tokenizing on either words or characters.
+The following generated samples were created using character-tokenized sequences.
+Note that any capitalization is \textit{not} added by the author---it was generated by the model.
+Also note that the documentation stated to train the model for a mere five epochs.
+
+        \begin{quote}
+            ``@realDonaldTrump you would be a clue and it's a great season.
+The Second and a state and all why are you all there, a real estate of him right on his interests and entertaining the U.S.
+and the world and the great policies are a fantastic game!''
+
+            ``I will be on @foxandfriends at: A.M. Enjoy!''
+
+            ``The Trade Resort is killing our country picks at the borders that we can energy the president of a total leader and a winner of the most broke our since I all television clifful in the other politics of the borders are a great complete disaster.''
+
+            ``@realDonaldTrump the many of the Republican candidates and have been a big problem. \#Trump''
+
+            ``@realDonaldTrump @FoxNews  You are the only one that are with your badly because you support it.
+The man at our country will be great to have the mother is one of the big millions of comments are looking like a big people of the U.S.
+he will be so much more than he is the U.S.''
+
+            ``Highly improvement on back with Democrats of the coming evenings than the lands showication and one of the place tomorrow at: A.M.  Enjoy!''
+        \end{quote}
+
+        These results seem to be more coherent, but it does seem like it simply strings together substrings from the larger corpus.
+
 \section{Discussion}
-    \todoinline{
-        \begin{itemize}
-            \item How good was it?
-            \item Accuracy meaningless because attempt is creative generation.
-            \item Did it pick up on patterns backed by data?
-        \end{itemize}
-    }
+    Even taking the entirety of President Trump's tweets (including time not spent in office), cherry-picked results have a quite political flavor.
+This was unexpected, but welcome as it exacerbated the patterns shown in the generated corpus.
+Even so, it is difficult to get a good objective grasp of how well generational models perform due to the qualitative nature of their outputs.
+Therefore, accuracy analysis is essentially meaningless, especially because the intent of generational models is to \textit{avoid} exactly replicating instances in the training set.
 
-\section{Future Work}
-    \todoinline{
-        \begin{itemize}
-            \item Word level generation. Instead of each character being mapped to an integer, create a vocab list and map each word to an integer.
-            \item Word association norms?
-            \item Embeddings?
-            \item Combining sources?
-            \item Grammar and spelling, part of speech tagging, word association, linguistic features as a component of a GAN
-            \item Crowd-sourced adversarial network as a training component?
-        \end{itemize}
-    }
+    We ultimately recommend using the six-layer, 64 unit LSTM network illustrated in \autoref{fig:architecture-64x6} because the generated output seemed to be more diverse, with less repetition.
+However, we do have recommendations for future work in this area.
+There has been an enormous amount of recent research in natural language understanding using machine learning models.
+The author suggest any further research look into using spaCy \citep{spacy2} to strip each word to its root form, perform part-of-speech tagging, grammar analysis, word association analysis, and other linguistic analysis as a one or more components of a Generative Adversarial Network (GAN).
+
+    The best performing models as a whole generated fairly reasonable results, but as \citet{karpathy_2015} notes, most recurrent generational models produce results that must be scraped by hand to find the best results.
+We recommend a different strategy for future work however.
+It seems like the Textgenrnn model could be used as the generative portion of a Generative Adversarial Network, and use some kind of grammar metric as the adversarial component of the model.
+
+    Further, the author also suggests tokenizing at a word level so that the recurrent model predicts the next word in a sequence rather than the next character.
+This is drastically increase the size of the output vector, but will diminish the size of dataset.
+It will probably be necessary to prune out the least common words, as well as treat punctuation as individual tokens.
+
+    It would also be possible, if training time is not an issue, to crowd-source the adversarial component of a GAN by tweeting each generated tweet, and recording how many likes or retweets each received.
 

### 2018-04-24 experiments/dataset.py
Update project documentation

@@ -18,12 +18,17 @@ from zipfile import ZipFile
 import numpy as np
 
 
+# The symbol set the preprocessed dataset will use
+ALPHABET = frozenset(string.ascii_letters + string.punctuation + ' #@')
+
+
 def remove_urls(vTEXT):
     """
         Remove URLS from a given string.
     """
     vTEXT = re.sub(r'(https|http)?:\/\/(\w|\.|\/|\?|\=|\&|\%)*\b', '', vTEXT, flags=re.MULTILINE)
     return vTEXT
+
 
 def remove_mentions(vTEXT):
     """
@@ -33,18 +38,17 @@ def remove_mentions(vTEXT):
     return vTEXT
 
 
-ALPHABET = frozenset(string.ascii_letters + string.punctuation + ' #@')
-
-
 def preprocess(text, alphabet):
     """
-        Preprocess text using the frozenset `alphabet` given.
+        Preprocess text using the frozenset `alphabet` given. Will convert everything
+        to lower case.
     """
     return filter(alphabet.__contains__, text.lower())
 
+
 def load_dataset(files, verbose=True):
     """
-        Parses the tweets out of the zipped dataset. Returns a generator of preprocessed tweets.
+        Parses the tweets out of the zipped dataset. Returns an iterable of preprocessed tweets.
     """
     dataset = []
 
@@ -58,8 +62,8 @@ def load_dataset(files, verbose=True):
                     dataset += json.load(f)
 
     # Filter out the retweets
-    dataset = filter(lambda x: x['is_retweet'] is False, dataset)
-    # Get only the text
+    dataset = filter(lambda t: t['is_retweet'] is False, dataset)
+    # Get only the tweet text
     dataset = (t['text']for t in dataset)
     # Remove URLs from the tweets.
     dataset = (remove_urls(t) for t in dataset)
@@ -67,7 +71,6 @@ def load_dataset(files, verbose=True):
     # dataset = (remove_mentions(t) for t in dataset)
     # Preprocess each tweet, filtering out nonascii alphabetic
     dataset = [''.join(preprocess(t, ALPHABET)) for t in dataset]
-
     return dataset
 
 
@@ -75,7 +78,6 @@ def chunk_text(corpus, length, step):
     """
         Cut the given corpus into a series of offset sequences.
     """
-
     # A list of sequences
     sequences = []
     # For each sequence, record the next character in the corpus
@@ -84,6 +86,30 @@ def chunk_text(corpus, length, step):
[…]

### 2018-04-24 experiments/generate.py

@@ -13,15 +13,19 @@ from dataset import load_dataset, seq_data
 from model import BASENAME, DIVERSITIES, GEN_LEN, SEQ_LEN, SEQ_STEP, load_model
 
 
-def sample_predictions(preds, temperature=1.0):
+def sample_predictions(preds, diversity=1.0):
     """
-        Helper function to sample an index from a probability array.
+        Helper function to sample an index from a probability array
+        with some diversity. The diversity is so that we don't simply
+        take the largest probability. This is so we get some amount
+        of "creativity".
     """
     preds = np.asarray(preds).astype('float64')
-    preds = np.log(preds) / temperature
+    preds = np.log(preds) / diversity
     exp_preds = np.exp(preds)
     preds = exp_preds / np.sum(exp_preds)
     probas = np.random.multinomial(1, preds, 1)
+    # Return the index of the largest element
     return np.argmax(probas)
 
 
@@ -33,7 +37,7 @@ def generate_sequence(model, seed, length, diversities,
     g = []
     for diversity in diversities:
         generated = ''
-        sequence = seed
+        # Avoid saving the seed as "generated" text for easier parsing, reading, and interpretation
         # generated += seed
 
         if verbose:
@@ -44,15 +48,15 @@ def generate_sequence(model, seed, length, diversities,
 
         for _ in range(length):
             X_pred = np.zeros((1, seq_length, num_chars))
-            for t, c in enumerate(sequence):
+            for t, c in enumerate(seed):
                 X_pred[0, t, char_to_indices[c]] = 1
 
             predictions = model.predict(X_pred, verbose=0)[0]
             next_index = sample_predictions(predictions, diversity)
             next_char = indices_to_char[next_index]
             generated += next_char
-            # TODO: Don't use a list for this, use a queue
-            sequence = sequence[1:] + next_char
+            # Move the seed down one character to generate the next character
+            seed = seed[1:] + next_char
         g.append(generated)
         if verbose:
             print(generated)
@@ -61,7 +65,7 @@ def generate_sequence(model, seed, length, diversities,
 
 def save_generated(generated, filename):
     """
-        Saves the generated text strings to the given file.
+        Saves the generated text to the given (extensionless) filename.
     """
     with open(f'{filename}.txt', 'a') as f:
[…]

### 2018-04-24 experiments/model.py

@@ -10,8 +10,6 @@ from keras.layers import LSTM, Dense, BatchNormalization
 # from keras.layers import CuDNNLSTM as LSTM
 from keras.models import Sequential, model_from_json
 
-# from textgenrnn.textgenrnn import textgenrnn
-
 # Turn the corpus into sequences 60 characters long
 SEQ_LEN = 60
 # Make each sequence offset from each other by 3 characters
@@ -36,7 +34,9 @@ def build_model(seq_length, num_chars, verbose):
     """
     if verbose:
         print('Building model...')
+
     model = Sequential()
+
     model.add(LSTM(units=64, return_sequences=True, input_shape=(seq_length, num_chars), unit_forget_bias=True))
     model.add(BatchNormalization())
     model.add(LSTM(units=64, return_sequences=True, unit_forget_bias=True))
@@ -52,7 +52,6 @@ def build_model(seq_length, num_chars, verbose):
     model.add(Dense(num_chars, activation='softmax'))
 
     model.compile(loss='categorical_crossentropy', optimizer='rmsprop')
-    # model = textgenrnn(name='trumperator')
     return model
 
 

### 2018-04-24 experiments/train.py

@@ -13,7 +13,7 @@ import matplotlib
 
 # Hack to enable matplotlib to save a figure without an X server running (SSH sessions)
 # https://stackoverflow.com/a/4706614
-# Needs to be ran before `import matplotlib.pyplot as plt`
+# Needs to run before `import matplotlib.pyplot as plt`
 matplotlib.use('Agg')
 
 import matplotlib.pyplot as plt
@@ -25,9 +25,9 @@ from generate import generate
 from model import (BASENAME, BATCH_SIZE, EPOCHS, PERCENT_VALIDATION,
                    SEQ_LEN, SEQ_STEP, build_model, save_model)
 
-# from textgenrnn.textgenrnn import textgenrnn
+from textgenrnn.textgenrnn import textgenrnn
 
-# Hack to keep keras from allocating the whole damn gpu.
+# Hack to keep Keras from allocating the whole damn gpu.
 config = tf.ConfigProto()
 config.gpu_options.allow_growth = True
 config.gpu_options.visible_device_list = "0"
@@ -85,6 +85,7 @@ def main(verbose):
 
     # The data is shuffled so the validation data isn't simply the latest 20% of tweets
     X, y = vec_data(sequences, next_chars, SEQ_LEN, nc, c2i, verbose)
+    # Split off the last 20% as validation data for pretty graphs
     n = len(X)
     num_val = int(PERCENT_VALIDATION * n)
     X_val = X[n - num_val:]
@@ -99,10 +100,13 @@ def main(verbose):
     model = build_model(SEQ_LEN, nc, verbose)
     history = train_model(model, X_train, y_train, X_val, y_val, verbose)
     plot_model_loss(BASENAME, history, verbose)
+    # Save the trained model so we don't have to wait 25 hours to generate another 10 tweet sample
     save_model(model, BASENAME, verbose)
     # Generate sample tweets using 10 random seeds from the corpus.
     generate(BASENAME, model, corpus, c2i, i2c, nc, 10, verbose)
 
+    # # Train the Textgenrnn model on our dataset. The README has information on how to take
+    # # a trained model and generate tweets. (It's three lines of code)
     # gen = textgenrnn()
     # gen.train_new_model(
     #     dataset,
